\huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation

### 2019-08-07 Manuscript_Publication.Rmd

---
title: \huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation
  in birds}
author:
- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
  and Iliana Medina Guzmán\textsuperscript{§}
- '*\textit{justin.g.cally@gmail.com}'
bibliography: References/SS_Extinction.bib
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
  \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
  \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption}
  \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
  \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip}
  \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
  \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{}
  \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
  \raggedbottom
csl: References/evolution.csl
subparagraph: yes
abstract: Sexual selection is thought to shape phylogenetic diversity by affecting speciation or extinction rates. However, the net effect of sexual selection on diversification is hard to predict _a priori_, because many of the hypothesised effects on speciation or extinction have opposing signs and uncertain magnitudes. Here, we test whether the strength of sexual selection predicts variation in speciation and extinction rates across passerine birds (up to 5,812 species, covering most genera). We tested for associations between two measures of sexual selection (sexual dichromatism, plus an index that captures variance in sexual size dimorphism, polygyny, and male parental care), several environmental variables, and multiple measures of speciation and extinction rates. Our results show that male-biased sexual selection, but not sexual dichromatism, predicts speciation rates ($\lambda_{DR}$) in passerines, and found no evidence that this relationship varies with the environment. We also found a strong negative relationship between range size and speciation rate. There was no correlation between sexual selection and extinction rate **XXXXXX**. Our findings support the view that sexual selection is a potent evolutionary force that has shaped diversification in songbirds. \vspace{5mm}  \par\noindent
  \textbf{Keywords:} Speciation rate, Comparative analysis, Plumage dichromatism, Range size, Extinction rate.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

@@

<!-- Luke general comments:
In a few places you say things like "1000 random trees". To me this means trees with a random topology (which are sometimes used as some sort of null model), but I think you actually mean 1000 trees randomly sampled (without replacement) from the posterior of trees in Jetz et al. Is that right? If so, probably best to search for the word "random" and remove/reword it.

@@

Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and in most species it predominently operates on males [@Kokko_Jennions2008; @Fromhage_2016; @Janicke_2016]. There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and phylogenetic diversification [e.g. @Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018]. Recent empirical and methodological advances, such as growing evidence that sexual selection can profoundly affect many traits [e.g. life history, body mass, and immunocompetence; @Cally_2019], larger and more accurate phylogenies [@Jetz_2012], and new phylogenetic methods (**REFERENCES**), present new opportunities to test whether and how sexual selection drives diversification.

Several hypotheses predict a positive relationship across species between the strength of sexual selection and the rate of speciation and/or extinction. For example, sexual selection is involved in reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation (**REF**). Additionally, sexual selection often operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia]. Furthermore, sexual selection has been hypothesised to increase diversification by maintaining trait combinations that would be selected out under pure natural selection [@Bonduriansky_2011; @Radwan_2016], and by helping to prevent extinction by purging deleterious mutations [@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adapation in environments [@Lorch_2003; @Candolin_2008; @Cally_2019]. Also, sexual selection sometimes favours individuals with rare phenotypes [@holman2013crozier], providing a source of diversifying selection that might increase the rate at which isolated populations diverge genetically.

Conversely, there are multiple ways in which sexual selection might hinder speciation or make extinction more likely. Firstly, sexual selection might promote extinction by favouring traits that improve mating success but reduce population fitness. For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the reproductive success of individuals while reducing population-wide productivity [e.g. infanticide, harassment, and reduced parental care by mate-seeking males; @RN37; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016]. Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013]. Secondly, sexual selection might promote evolutionary stasis by ensuring that individuals with novel trait values have low mating success.

Importantly, the relationship between sexual selection and diversification might be contingent on the environment. Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [e.g. @Long_2012; @Connallon_Hall_2016], though the empirical evidence for this theory is mixed and largely limited to insects [@Cally_2019]. The mechanism underpinning this prediction is that the environment and the selective history of the population should influence the relative amounts of genetic variation at sexually concordant loci (*i.e.* loci where the same allele is fittest for both sexes) and sexually antagonistic loci (where different alleles are favoured in each sex). In stable environments, variation is thought to be preferentially depleted at sexually concordant loci, leading to stronger gender load and reduced net benefits of sexual selection [e.g. @Connallon_Hall_2016], and potentially to a more negative relationship between sexual selection and diversification rates across species. Conversely, sexual selection might be especially important at promoting local adaptation when the environment displays variability at the relevant spatial and temporal scales [e.g. @Boughman_2002; @Lorch_2003; @Maan_2011; @kokko2008condition]

<!-- .; causing displacement of male and female traits from their optimum. It follows from these predictions that populations with stronger sexual selection would have fitness benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008]. Additionally, sexual selection can reinforce local adaptations through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011]. For instance, beak morphology is a trait that arose under natural selection in several taxa and is now a trait under sexual selection [a.k.a. *magic traits*; @Servedio_2011]. In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007]. Through these synergetic effects sexual selection and natural selection (*e.g.* environmental variation over space and time) may lead to increased speciation rates.   -->

Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [*e.g.* @Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014]. Birds are well-suited for this purpose due to their diverse and well-characterised mating systems and sexually-selected traits, and because their phylogenetic relationships are comparatively well-known. A 2011 meta-analysis covering 20 primary studies of birds and other taxa found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects or fish [@Kraaijeveld_2011]. However, there was large variation in effect size estimates across the 20 studies, likely reflecting differences in methodology, such as metrics used to characterise speciation and sexual selection, in addition to true biological differences. More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012]. Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.

Here, we investigate the association between sexual selection and diversification in birds while building upon previous approaches in multiple ways. We use two measures of the strength of sexual selection: sexual dichromatism [*n* = 5,812 species; estimated from illustrations by @Dale_2015], as well as an index of male-biased sexual selection [*n* = 2,465 species; derived via principal components analysis by @Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care. Most **OR ALL??** earlier macroevolutionary studies of this topic only quantified sexual selection using dichromatism, which is problematic since dichromatism does not always signal the presence of strong sexual selection and _vice versa_ [@Dale_2015]. For example, male and female dunnocks (_Prunella modularis_) are similarly coloured yet sexual selection appears to be strong [@davies1986re]. Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018]. Additionally, our analysis includes multiple ecological and environmental variables on diversification, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016]. We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$). Finally, we incorporate phylogenetic uncertainty into our analyses, improving over earlier work which assumed the statistically best-fitting phylogeny is correct.  \phantomsection

@@

**[Is it possible to move this later in the Results? It's possibly not ideal to start by highlighting the dependence of our conclusions on the methods we choose, and not mentioning our main non-null finding regarding the SS index. Having said that I can see a case for leaving it here too. - Luke]**

We first estimated speciation rates across the phylogeny, using the maximum clade credibility (MCC) tree, as well as 100 or 1000 trees sampled from the posterior distribution of trees in @Jetz_2012. Estimates of speciation rates varied between phylogenetic trees, especially for rates calculated by BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95% highest posterior density (HPD) interval across 100 trees was >20-fold larger than the **95% confidence interval [SHOULD THIS BE HPD TOO?]** calculated for the MCC tree. Conversely, for our estimates of diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$), the 95% HPD interval across 1,000 trees was near-identical to the 95% confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}). Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable. Mean measures of speciation rate across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}). Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}). Full details of the BAMM results are presented as supplementary materials.

@@

**[I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Abstract and Intro. - Luke]**

We next fit Phylogenetic Generalized Least Squares (PGLS) models with **2-way [CORRECT?]** interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) and compared them using AICc model selection. However, no interaction terms were present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}). Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation. Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).   
\phantomsection

@@

We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation. Using three different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) as the response variable, the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree. Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}). Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 trees for $\lambda_{BAMM}$. The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}). For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).  \phantomsection

@@

We would like to acknowledge Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees. **Any others?**

>>>>

---
title: \huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation
  in birds}
author:
- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§}
  and Iliana Medina\textsuperscript{§}
- '*\textit{justin.g.cally@gmail.com}'
bibliography: References/SS_Extinction.bib
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
  \rhead[]{Sexual selection predicts speciation} \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
  \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption}
  \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
  \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip}
  \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
  \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{}
  \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{}
  \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang,
  flushmargin]{footmisc} \usepackage{hyperref}\hypersetup{colorlinks=true, hyperfootnotes=false,
  linkcolor=blue, filecolor=blue, urlcolor=blue} \usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}
  \raggedbottom
csl: References/NComms.csl
subparagraph: yes
abstract: Sexual selection is thought to shape phylogenetic diversity by affecting
  speciation or extinction rates. However, the net effect of sexual selection on diversification
  is hard to predict _a priori_, because many of the hypothesised effects on speciation
  or extinction have opposing signs and uncertain magnitudes. Here, we test whether
  the strength of sexual selection predicts variation in speciation and extinction
  rates across passerine birds (up to 5,812 species, covering most genera). We tested
  for associations between two measures of sexual selection (sexual dichromatism and
  a multivariate measure of male-biased sexual selection), several environmental variables,
  and multiple measures of speciation and extinction rates. Our results show that
  male-biased sexual selection, but not sexual dichromatism, predicts speciation rates
  ($\lambda_{DR}$) in passerines, and found no evidence that this relationship varies
  with the environment. We also found a strong negative relationship between range
  size and speciation rate. There was no correlation between sexual selection and
  extinction rate ($\mu_{BAMM}$). Our findings show that sexual dichromatism is a
  poor proxy for sexual selection at least a broad-scale and support the view that
  sexual selection is a potent evolutionary force that has shaped diversification
  analysis, Plumage dichromatism, Range size, Extinction rate.
  in songbirds. \vspace{5mm}  \par\noindent \textbf{Keywords:} Speciation rate, Comparative
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia}

@@

<!-- Luke general comments:
In a few places you say things like "1000 rrandom trees". To me this means trees with a random topology (which are sometimes used as some sort of null model), but I think you actually mean 1000 trees randomly sampled (without replacement) from the posterior of trees in Jetz et al. Is that right? If so, probably best to search for the word "random" and remove/reword it.

@@

Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and in most species it predominently operates on males [@Kokko_Jennions2008; @Fromhage_2016; @Janicke_2016]. There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and phylogenetic diversification [e.g. @Lande_1981; @Lande_1982; @west1983sexual; @Seddon_2008; @Cooney_2018]. Recent empirical and methodological advances, such as growing evidence that sexual selection can profoundly affect many traits [e.g. life history, body mass, and immunocompetence; @Cally_2019], larger and more accurate phylogenies [@Jetz_2012], and new phylogenetic methods [@Rabosky_2014; @Harvey_essim_2017], present new opportunities to test whether and how sexual selection drives diversification.

Several hypotheses predict a positive relationship across species between the strength of sexual selection and the rate of speciation and/or extinction. For example, sexual selection is involved in reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation [see @Ritchie_2007]. Additionally, sexual selection often operates on traits that can create reproductive isolation when they diverge between lineages, such as signals and preferences involved in mate selection [@Lande_1981; @Lande_1982; @Safran_2013], sperm-egg interactions [@swanson1998abalone], or genital morphology [@sloan2019genitalia]. Furthermore, sexual selection has been hypothesised to increase diversification by maintaining trait combinations that would be selected out under pure natural selection [@Bonduriansky_2011; @Radwan_2016], and by helping to prevent extinction by purging deleterious mutations [@Whitlock_Agrawal_2009], fixing beneficial ones [@Whitlock_2000], and accelerating adapation in environments [@Lorch_2003; @Candolin_2008; @Cally_2019].

Conversely, sexual selection might hinder speciation or make extinction more likely by favouring traits that improve mating success but reduce population fitness. For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the reproductive success of individuals while reducing population-wide productivity [e.g., infanticide, harassment, and reduced parental care by mate-seeking males; @RN37; @Kokko_Jennions2008; @Holman_Kokko_2013; @Fromhage_2016]. Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @harano2010intralocus; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013].

Importantly, the relationship between sexual selection and diversification might be contingent on the environment. Theoretical work predicts that sexual selection should have a more positive effect on adaptation and population fitness in variable environments relative to stable ones [e.g. @Long_2012; @Connallon_Hall_2016], though the empirical evidence for this theory is mixed and largely limited to insects [@Cally_2019]. The mechanism underpinning this prediction is that the environment and the selective history of the population should influence the relative amounts of genetic variation at sexually concordant loci (i.e. loci where the same allele is fittest for both sexes) and sexually antagonistic loci (where different alleles are favoured in each sex). In stable environments, variation is thought to be preferentially depleted at sexually concordant loci, leading to stronger gender load and reduced net benefits of sexual selection [e.g. @Connallon_Hall_2016], and potentially to a more negative relationship between sexual selection and diversification rates across species. Conversely, sexual selection might be especially important at promoting local adaptation when the environment displays variability at the relevant spatial and temporal scales [e.g. @Boughman_2002; @Lorch_2003; @Maan_2011; @kokko2008condition]

Birds have been a popular focus of macroevolutionary studies of sexual selection and diversification [e.g. @Barraclough_1995; @Morrow_Speciation_2003; @Seddon_2008; @Seddon_2013; @Huang_2014]. Birds are well-suited for this purpose due to their diverse and well-characterised mating systems and sexually-selected traits, and because their phylogenetic relationships are comparatively well-known. A 2011 meta-analysis covering 20 primary studies of birds and other taxa found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects or fish [@Kraaijeveld_2011]. However, there was large variation in effect size estimates across the 20 studies, likely reflecting differences in methodology, such as metrics used to characterise speciation and sexual selection, in addition to true biological differences. More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012]. Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores provided by human observers.

Here, we investigate the association between sexual selection and diversification in birds while building upon previous approaches in multiple ways. We use two measures of the strength of sexual selection: sexual dichromatism [*n* = 5,812 species; estimated from illustrations by @Dale_2015], as well as an index of male-biased sexual selection [*n* = 2,465 species; derived via principal components analysis by @Dale_2015], which captures (co)variation in sexual size dimorphism, social polygyny and paternal care. These two measures are used because sexual dichromatism does not always signal the presence of strong sexual selection and _vice versa_ [@Dale_2015]. For example, male and female dunnocks (_Prunella modularis_) are similarly coloured yet sexual selection appears to be strong [@davies1986re]. Furthermore, a recent comparative study found a negative relationship between dichromatism and another sexually-selected trait (song) across species, suggesting that a multi-trait focus would improve estimates of sexual selection intensity [@Cooney_2018]. Additionally, our analysis includes multiple ecological and environmental variables on diversification, allowing us to adjust for potential confounds, to identify environmental factors driving diversification, and to test whether environmental factors interact with sexual selection as theory predicts [@Connallon_Hall_2016]. We use multiple approaches for quantifying speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; see, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).  \phantomsection

@@

**[Is it possible to move this later in the Results? It's possibly not ideal to start by highlighting the dependence of our conclusions on the methods we choose, and not mentioning our main non-null finding regarding the SS index. Having said that I can see a case for leaving it here too. - Luke] [We initially had this at the end but moved it up to the start. What do you think Devi? - Justin]**

Estimates of the effect of predictor variables on speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % highest posterior density (HPD) interval across phylogenetic least squares (PGLS) model estimates from 100 trees was often > 20 times larger than the 95 % confidence interval for estimates from a single PGLS model using the MCC (maximum clade credibility) tree. This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval of model estimates for pgls models using 1,000 trees was near-equivelant to the 95 % confidence interval calculated for pgls model estimates of the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}). Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar statistical support is notable. Mean measures of speciation rate across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}). Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}). Full details of the BAMM results are presented as supplementary materials.

@@

**[I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Abstract and Intro. - Luke] [Statistically, I think it is best to rule out the interactions first but Yes I see your point - Justin]**

We next fit PGLS models with 2-way interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) were included in the full model for posterior model selection. However, no interaction terms were present in the top models ($\Delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\Delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}). Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation. Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).   
\phantomsection

@@

We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation. Using three different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) as the response variable, the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree. Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; i.e. no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}). Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 trees for $\lambda_{BAMM}$. The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top$SDi, format = "e", digits = 3)`, $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top$SDi, format = "e", digits = 3)`,\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}). For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).  \phantomsection

@@

We would like to acknowledge Adnan Moussalli for his assistance in conducting BAMM runs across 100 phylogenetic trees.
